Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-15 write-ups/PLDI/paper.tex
Beginning of subsection on performance and responsiveness

\ADComment{Preamble}

\HEComment{Below, just random notes so far}

Cooperative scheduling enables to concurrently run non-persistent
kernels alongside blocking ones. In terms of performance, we are
interested to see whether the variations of workgroups for the blocking
kernel.

Performance of cooperative scheduling corresponds to execution time of
cooperative kernels compared to the serial

In order to simulate non-persistent workloads alongside kernels of
blocking algorithms, we use a matrix multiplication kernel where
matrix dimensions are tuned to trigger the relevant execution
duration.

@@

\paragraph{Results}  \ADComment{Discuss the results.}

@@

There has been a lot of work on accelerating blocking irregular
algorithms using GPUs~\cite{...,TPO10}, and on the \emph{persistent threads}
programming style for long-running kernels~\cite{...}.  These
approaches rely on the occupancy-bound execution model, and aim to
take ownership of the GPU by scheduling as many workgroups as the GPU
has compute units, relying on fair scheduling between these
workgroups.  This means the GPU is unavailable for other tasks, and
the method is not future-proof because (as discussed throughout the
paper) the occupancy-bound execution model is not guaranteed by GPU
programming models and is unlikely to be guaranteed in practice by
future GPU platforms.

@@

Hardware support for preemptive multitasking for \nvidia GPUs has been
proposed, enabling full preemption as well as \emph{SM-draining},
whereby workgroups occupying a symmetric multiprocessor (SM; a compute
unit using our terminology) are drained until the SM becomes free for
other tasks.  A follow-up work adds the notion of SM \emph{flushing},
whereby a workgroup running on an SM can be cancelled and re-scheduled
from scratch if the workgroup has not yet committed any visible
side-effects to shared memory.  Both approaches have been evaluated
using simulators, over sets of regular GPU kernels.  State-of-the-art \nvidia GPUs targeting data centers (where multi-tasking is essential) do support
preemption, though it is not clear whether they guarantee fair scheduling~\cite{PascalWhitepaper}.

@@

\paragraph{Dynamic parallelism} Recent version of CUDA and OpenCL allow a kernel to spawn further kernels~\cite{CUDASpec,opencl2Spec}.
This \emph{dynamic parallelism} can be exploited to implement a
GPU-based scheduler, by having an initial scheduler kernel repeatedly
spawn further kernels as required, according to some scheduling
policy~\cite{DBLP:conf/ppopp/Muyan-OzcelikO16}.  However, a kernel
that uses dynamic parallelism is still subject to the scheduling
limitations of current GPU programming models, and in particular the
problem that workgroups are not fairly scheduled.  Hence this feature
does not offer a solution to the problem of deploying blocking
algorithms on GPUs in a safe manner that we tackle with this work.

>>>>

Cooperative scheduling enables to run a non-cooperative kernel alongside
a cooperative one, we focus here on the performance impact of this
multitasking and on the scheduling responsiveness. The performance is
measured with respect to the performance of each kernel run in
sequence. Responsiveness, i.e., how fast can a non-cooperative kernel
can be scheduled, depends not only on the scheduler implementation, but
also on how often the cooperative kernel offer workgroups to be killed.

% In order to simulate non-persistent workloads alongside kernels of
% blocking algorithms, we use a matrix multiplication kernel where
% matrix dimensions are tuned to trigger the relevant execution
% duration.

For a given cooperative kernel and its data set, we make it run while
regularly asking to schedule a non-cooperative kernel that simulates
graphics rendering, according to the classes listed in
\mytab~\ref{tab:noncooperativeworkload}. In practice, we use a matrix
multiplication kernel as the non-cooperative workload, with matrix
dimensions tailored for each GPU to reach the appropriate execution
duration when this non-cooperative kernel have access to all available
workgroups. Besides, we have different run where the non-cooperative
kernel is scheduled with different number of workgroups: either 1, a
quarter, a half or all but one workgroups among the ones available for
the cooperative kernel. Moreover, in graph algorithms we try both the
regular and the query barrier.

Our experiments span 13 pairs of cooperative kernels and data sets, 3
classes for the non-cooperative kernel, 4 quantities of workgroups
claimed for the non-cooperative kernel and 2 variations of resizing
barriers for graph algorithms, leading to 288 configurations.

A run is characterized by the cooperative kernel, its data set, the
non-cooperative kernel, its size, the number of workgroups we ask for
the non-cooperative kernel. This forms a configuration.

For each GPU, we first

@@

run cooperative kernels and fire non-cooperative kernels regularly,
asking for different number of workgroups

explain gather time, execution time. related to responsiveness and
performance

explain query barrier, or mention the results with / without query
barriers.

recall 10 runs for each configurations. How many configurations ?

Mention that some configuration failed.

We check correctness (completion) of both kernels. Some errors
sometimes.

\paragraph{Results}  \ADComment{Discuss the results.}
explain horizontal lines

missing results for all but one in bfs rmat22 since errors

also, high gather time corresponds to applications and data sets where
resize/kill is triggered less often.

Each configuration was run 10 times on each GPU.

@@

There has been a lot of work on accelerating blocking irregular
algorithms using GPUs~\cite{...}, and on the \emph{persistent threads}
programming style for long-running kernels~\cite{...}.  These
approaches rely on the occupancy-bound execution model, and aim to
take ownership of the GPU by scheduling as many workgroups as the GPU
has compute units, relying on fair scheduling between these
workgroups.  This means the GPU is unavailable for other tasks, and
the method is not future-proof because (as discussed throughout the
paper) the occupancy-bound execution model is not guaranteed by GPU
programming models and is unlikely to be guaranteed in practice by
future GPU platforms.

@@

Hardware support for preemptive multitasking for \nvidia GPUs has been
proposed, enabling full preemption as well as \emph{SM-draining},
whereby workgroups occupying a symmetric multiprocessor (SM; a compute
unit using our terminology) are drained until the SM becomes free for
other tasks.  A follow-up work adds the notion of SM \emph{flushing},
whereby a workgroup running on an SM can be cancelled and re-scheduled
from scratch if the workgroup has not yet committed any visible
side-effects to shared memory.  Both approaches have been evaluated
using simulators, over sets of regular GPU kernels.

@@

\paragraph{Dynamic parallelism} Recent version of CUDA and OpenCL allow a kernel to spawn further kernels~\cite{...}.
This \emph{dynamic parallelism} can be exploited to implement a GPU-based scheduler: an initial scheduler kernel uses the feature to spawn additional kernels as required~\cite{ }.

 is launched

%
\ADComment{Kepler preemption support?  Mention in related work?}

\ADComment{Sreepathi mentioned the NCCL framework.}
